Exploring how changes in cause of death have narrowed the black-white life expectancy gap in the United States.

### 2017-02-23 Manuscript/Manuscript_COD_decomp.Rmd

cod.df <- merge(cod.df, BlackWhite_results[BlackWhite_results$year == 1969 & BlackWhite_results$sex == "Male", 
                                           c("state", "Census_Region", "Census_Division", "stabbrs")], by = c("state"))
```

@@

  assign(paste0("max", i), max.value)
  assign(paste0("max.cause", i), max.cause)
  assign(paste0("min", i), min.value)
  assign(paste0("min.cause", i), min.cause)
}

cod.df.males <- merge(subset(cod.df, sex == "Male"), bw_dat.males, by = c("state"))
cod.df.males <- cod.df.males %>% mutate(new.start.1 = -start + first.gap, new.finish.1 = -finish + first.gap,
                                        new.start.2 = -start + second.gap, new.finish.2 = -finish + second.gap,
                                        new.start.3 = -start + third.gap, new.finish.3 = -finish + third.gap)

@@

**Fig 1: Decomposition of the change in the male black-white life expectancy gap by cause of death between 1969 and 1983 by US state**

```{r figure1, fig.align = "center", echo=F, fig.height=9, fig.width=7}
y.labels <- levels(factor(cod.df.males$state.gdiff.order.1))
y.num.breaks <- length(levels(factor(cod.df.males$state.gdiff.order.1)))

males.era1 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era1"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = Cause.of.death), 
              color = "white") + # + scale_alpha(range = c(0.64, 4.04) 
    geom_segment(data = subset(cod.df.males, Cause.of.death == "Cancers" & era == "era1"), aes(x = second.gap, xend = second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) + 
        geom_segment(data = subset(cod.df.males, Cause.of.death == "Cancers" & era == "era1"), aes(x = first.gap, xend = first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) + 
    theme_minimal() + 
    scale_x_continuous(breaks = seq(2, 16, by = 2), labels = seq(2, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

males.era1

plot <-   ggplot(subset(cod.df.males, era == "era1"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = Cause.of.death), #, alpha = tot_Cx_CI_width
              color = "white") + # + scale_alpha(range = c(0.64, 4.04) 
    geom_segment(data = subset(cod.df.males, Cause.of.death == "Cancers" & era == "era1"), aes(x = second.gap, xend = second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) + 
        geom_segment(data = subset(cod.df.males, Cause.of.death == "Cancers" & era == "era1"), aes(x = first.gap, xend = first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) + 
    theme_minimal() + 
    scale_x_continuous(breaks = seq(2, 16, by = 2), labels = seq(2, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)
```

@@

males.era2 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era2"), aes(y = state.gdiff.num.2, x = new.start.2)) +
    geom_rect(aes(xmin = new.start.2, ymin = state.gdiff.num.2 - 0.45, 
                  ymax = state.gdiff.num.2 + 0.45, 
                  xmax = new.finish.2, fill = Cause.of.death), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5), lty = 3) +
    theme_minimal() + 
    scale_x_continuous(breaks = seq(2, 16, by = 2), labels = seq(2, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks.2, labels = y.labels.2), minor_breaks = NULL) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

@@

males.era3 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era3"), aes(y = state.gdiff.num.3, x = new.start.3)) +
    geom_rect(aes(xmin = new.start.3, ymin = state.gdiff.num.3 - 0.45, 
                  ymax = state.gdiff.num.3 + 0.45, 
                  xmax = new.finish.3, fill = Cause.of.death), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = fourth.gap, xend= fourth.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5), lty = 3) +
    theme_minimal() + 
    scale_x_continuous(breaks = seq(0, 16, by = 2), labels = seq(0, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks.3, labels = y.labels.3)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1993-2013"), yaxis = list(title = NA, autorange = "reversed"))

@@

```{r, fig.align="center", fig.height=5, fig.width=9, echo=F, eval = F}
males.era2.2 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era2"), aes(y = state.gdiff.num.1, x = new.start.2)) +
    geom_rect(aes(xmin = new.start.2, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.2, fill = Cause.of.death), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

@@

subplot(males.era1, males.era2.2, shareY = T)
```
** **
** **

@@

cod.df.females <- merge(subset(cod.df, sex == "Female"), bw_dat.females, by = c("state"))
cod.df.females <- cod.df.females %>% mutate(new.start.1 = -start + first.gap, new.finish.1 = -finish + first.gap,
                                        new.start.2 = -start + second.gap, new.finish.2 = -finish + second.gap,
                                        new.start.3 = -start + third.gap, new.finish.3 = -finish + third.gap)

@@

females.era1 <- ggplotly(
  ggplot(subset(cod.df.females, era == "era1"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = Cause.of.death), 
              color = "white") +
    geom_segment(data = bw_dat.females,
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.females,
                 aes(x = first.gap, xend= first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

@@

females.era2 <- ggplotly(
  ggplot(subset(cod.df.females, era == "era2"), aes(y = state.gdiff.num.2, x = new.start.2)) +
    geom_rect(aes(xmin = new.start.2, ymin = state.gdiff.num.2 - 0.45, 
                  ymax = state.gdiff.num.2 + 0.45, 
                  xmax = new.finish.2, fill = Cause.of.death), 
              color = "white") +
    geom_segment(data = bw_dat.females,
                 aes(x = third.gap, xend = third.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5)) +
    geom_segment(data = bw_dat.females,
                 aes(x = second.gap, xend = second.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks.2, labels = y.labels.2)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1983-1993"), yaxis = list(title = NA, autorange = "reversed"))

@@

females.era3 <- ggplotly(
  ggplot(subset(cod.df.females, era == "era3"), aes(y = state.gdiff.num.3, x = new.start.3)) +
    geom_rect(aes(xmin = new.start.3, ymin = state.gdiff.num.3 - 0.45, 
                  ymax = state.gdiff.num.3 + 0.45, 
                  xmax = new.finish.3, fill = Cause.of.death), 
              color = "white") +
    geom_segment(data = bw_dat.females,
                 aes(x = fourth.gap, xend = fourth.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5)) +
    geom_segment(data = bw_dat.females,
                 aes(x = third.gap, xend = third.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks.3, labels = y.labels.3)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1993-2013"), yaxis = list(title = NA, autorange = "reversed"))

@@

| Broad cause of death     | Included causes of death                                                     |
|--------------------------|------------------------------------------------------------------------------|
| Cancers                   | Lip; Tongue; Salivary gland; Floor of mouth; Gum and other mouth; Nasopharnyx; Tonsil; Oropharynx; Hypopharynx; Other oral cavity and pharynx; Esophagus; Stomach; Small intestine; Colon excluding rectum; Rectum and rectosigmoid junction; Anus, anal canal and anorectum; Liver; Intrahepatic bile duct; Gallbladder; Other biliary; Pancreas; Retroperitoneum; Peritoneum, Omentum and mesentery; Other digestive organs; Nose, nasal cavity and middle ear; Larynx; Lung and bronchus; Pleura; Trachea, Mediastinum and other respiratory organs; Bones and joints; Soft tissue including heart; Melanoma of the skin; Non-melanoma skin; Breast; Cervix uteri; Corpus uteri; Uterus, NOS; Ovary; Vagina; Vulva; Other female genital organs; Prostate; Testis; Penis; Other male genital organs; Urinary bladder; Kidney and renal pelvis; Ureter; Other urinary organs; Eye and orbit; Brain and other nervous system; Thyroid; Other endocrine including thymus; Hodkin lymphoma; Myeloma; Acute lymphocytic leukemia; Chronic lymphocytic leukemia; Other lymphocytic leukemia; Acute myeloid leukemia; Acute monocytic leukemia; Chronic myeloid leukemia; Other myeliod/monocytic leukemia; Other acute leukemia; Aleukemic, Subleukemic and NOS; Miscellaneous malignant cancer. |
| Cardiovascular disease   | Diseases of heart; Hypertension without heart disease; Cerebrovascular diseases; Atherosclerosis; Aortic aneurysm and dissection; Other diseases of arteries, arterioles, capillaries. | 
| Communicable disease     | Tuberculosis; Syphilis; HIV (1987+), Septicemia; Other infectious and parasitic diseases; Pneumonia and influenza |
| Non-communicable disease | Diabetes Mellitus; Alzheimers (ICD9 and 10 only); Chronic obstructive pulmonary disease and allied conditions; Stomach and duodenal ulcers; Chronic liver disease and cirrhosis; Nephritis, Nephrotic Sydrome and Nephrosis.  |
| Injuries                 | Accidents and Adverse Effects; Suicide and Self-inflicted Injury; Homicide and Legal Intervention. |
| All other causes         | Complications of pregnancy, childbirth, puerperium; Congenital anomalies; Certain conditions originating in perinatal period; Symptoms, signs, and ill-defined conditions; Other cause of death. |

>>>>

cod.df <- merge(cod.df, BlackWhite_results[BlackWhite_results$year == 1969 & BlackWhite_results$sex == "Male", 
                                           c("state", "Census_Region", "Census_Division", "stabbrs")], by = c("state"))
cod.df.males <- merge(subset(cod.df, sex == "Male"), bw_dat.males, by = c("state"))
cod.df.males <- cod.df.males %>% mutate(new.start.1 = -start + first.gap, new.finish.1 = -finish + first.gap,
                                        new.start.2 = -start + second.gap, new.finish.2 = -finish + second.gap,
                                        new.start.3 = -start + third.gap, new.finish.3 = -finish + third.gap)

cod.df.males$tot_Cx_CI_width2 <- -(cod.df.males$tot_Cx_CI_width - min(cod.df.males$tot_Cx_CI_width))/(max(cod.df.males$tot_Cx_CI_width) - min(cod.df.males$tot_Cx_CI_width)) + 1
cod.df.males$width.factor <- as.factor(cut(cod.df.males$tot_Cx_CI_width2, breaks = c(-0.01, 0.25, 0.50, 0.75, 1)))

@@

  assign(paste0("max", i), max.value)
  assign(paste0("max.cause", i), max.cause)
  assign(paste0("min", i), min.value)
  assign(paste0("min.cause", i), min.cause)
}
```

@@

**Fig 1: Decomposition of the change in the male black-white life expectancy gap by cause of death between 1969 and 1983 by US state**
```{r figure1, fig.align = "center", echo=F, fig.height=9, fig.width=7}
y.labels <- levels(factor(cod.df.males$state.gdiff.order.1))
y.num.breaks <- length(levels(factor(cod.df.males$state.gdiff.order.1)))

males.era1 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era1"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = Cause.of.death, alpha = width.factor), 
              color = "white") +
    geom_segment(data = subset(cod.df.males, Cause.of.death == "Cancers" & era == "era1"), aes(x = second.gap, xend = second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) + 
        geom_segment(data = subset(cod.df.males, Cause.of.death == "Cancers" & era == "era1"), aes(x = first.gap, xend = first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) + 
    theme_minimal() + 
    scale_x_continuous(breaks = seq(2, 16, by = 2), labels = seq(2, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

males.era1
```

@@

males.era2 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era2"), aes(y = state.gdiff.num.2, x = new.start.2)) +
    geom_rect(aes(xmin = new.start.2, ymin = state.gdiff.num.2 - 0.45, 
                  ymax = state.gdiff.num.2 + 0.45, 
                  xmax = new.finish.2, fill = Cause.of.death, alpha = width.factor), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5), lty = 3) +
    theme_minimal() + 
    scale_x_continuous(breaks = seq(2, 16, by = 2), labels = seq(2, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks.2, labels = y.labels.2), minor_breaks = NULL) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

@@

males.era3 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era3"), aes(y = state.gdiff.num.3, x = new.start.3)) +
    geom_rect(aes(xmin = new.start.3, ymin = state.gdiff.num.3 - 0.45, 
                  ymax = state.gdiff.num.3 + 0.45, 
                  xmax = new.finish.3, fill = Cause.of.death, alpha = width.factor), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = fourth.gap, xend= fourth.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5), lty = 3) +
    theme_minimal() + 
    scale_x_continuous(breaks = seq(0, 16, by = 2), labels = seq(0, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks.3, labels = y.labels.3)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1993-2013"), yaxis = list(title = NA, autorange = "reversed"))

@@

```{r, fig.align="center", fig.height=5, fig.width=9, echo=F, eval = F}
males.era2.2 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era2"), aes(y = state.gdiff.num.1, x = new.start.2)) +
    geom_rect(aes(xmin = new.start.2, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.2, fill = Cause.of.death, alpha = width.factor), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

@@

males.era3.2 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era3"), aes(y = state.gdiff.num.1, x = new.start.3)) +
    geom_rect(aes(xmin = new.start.3, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.3, fill = Cause.of.death, alpha = width.factor), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = fourth.gap, xend= fourth.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

males.era3.2

subplot(males.era1, males.era2.2, males.era3.2, shareY = T)
```
** **
** **

@@

cod.df.females <- merge(subset(cod.df, sex == "Female"), bw_dat.females, by = c("state"))
cod.df.females$tot_Cx_CI_width2 <- -(cod.df.females$tot_Cx_CI_width - min(cod.df.females$tot_Cx_CI_width))/(max(cod.df.females$tot_Cx_CI_width) - min(cod.df.females$tot_Cx_CI_width)) + 1
cod.df.females$width.factor <- as.factor(cut(cod.df.females$tot_Cx_CI_width2, breaks = c(-0.01, 0.25, 0.50, 0.75, 1)))

cod.df.females <- cod.df.females %>% mutate(new.start.1 = -start + first.gap, new.finish.1 = -finish + first.gap,
                                        new.start.2 = -start + second.gap, new.finish.2 = -finish + second.gap,
                                        new.start.3 = -start + third.gap, new.finish.3 = -finish + third.gap)

@@

females.era1 <- ggplotly(
  ggplot(subset(cod.df.females, era == "era1"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = Cause.of.death, alpha = width.factor), 
              color = "white") +
    geom_segment(data = bw_dat.females,
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.females,
                 aes(x = first.gap, xend= first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

@@

females.era2 <- ggplotly(
  ggplot(subset(cod.df.females, era == "era2"), aes(y = state.gdiff.num.2, x = new.start.2)) +
    geom_rect(aes(xmin = new.start.2, ymin = state.gdiff.num.2 - 0.45, 
                  ymax = state.gdiff.num.2 + 0.45, 
                  xmax = new.finish.2, fill = Cause.of.death, alpha = width.factor), 
              color = "white") +
    geom_segment(data = bw_dat.females,
                 aes(x = third.gap, xend = third.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5)) +
    geom_segment(data = bw_dat.females,
                 aes(x = second.gap, xend = second.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks.2, labels = y.labels.2)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1983-1993"), yaxis = list(title = NA, autorange = "reversed"))

@@

females.era3 <- ggplotly(
  ggplot(subset(cod.df.females, era == "era3"), aes(y = state.gdiff.num.3, x = new.start.3)) +
    geom_rect(aes(xmin = new.start.3, ymin = state.gdiff.num.3 - 0.45, 
                  ymax = state.gdiff.num.3 + 0.45, 
                  xmax = new.finish.3, fill = Cause.of.death, alpha = width.factor), 
              color = "white") +
    geom_segment(data = bw_dat.females,
                 aes(x = fourth.gap, xend = fourth.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5)) +
    geom_segment(data = bw_dat.females,
                 aes(x = third.gap, xend = third.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks.3, labels = y.labels.3)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1993-2013"), yaxis = list(title = NA, autorange = "reversed"))

@@

| Broad cause of death     | Included causes of death                                                     |
|--------------------------|------------------------------------------------------------------------------|
| Cancers                   | Lip; Tongue; Salivary gland; Floor of mouth; Gum and other mouth; Nasopharnyx; Tonsil; Oropharynx; Hypopharynx; Other oral cavity and pharynx; Esophagus; Stomach; Small intestine; Colon excluding rectum; Rectum and rectosigmoid junction; Anus, anal canal and anorectum; Liver; Intrahepatic bile duct; Gallbladder; Other biliary; Pancreas; Retroperitoneum; Peritoneum, Omentum and mesentery; Other digestive organs; Nose, nasal cavity and middle ear; Larynx; Lung and bronchus; Pleura; Trachea, Mediastinum and other respiratory organs; Bones and joints; Soft tissue including heart; Melanoma of the skin; Non-melanoma skin; Breast; Cervix uteri; Corpus uteri; Uterus, NOS; Ovary; Vagina; Vulva; Other female genital organs; Prostate; Testis; Penis; Other male genital organs; Urinary bladder; Kidney and renal pelvis; Ureter; Other urinary organs; Eye and orbit; Brain and other nervous system; Thyroid; Other endocrine including thymus; Hodkin lymphoma; Myeloma; Acute lymphocytic leukemia; Chronic lymphocytic leukemia; Other lymphocytic leukemia; Acute myeloid leukemia; Acute monocytic leukemia; Chronic myeloid leukemia; Other myeliod/monocytic leukemia; Other acute leukemia; Aleukemic, Subleukemic and NOS; Miscellaneous malignant cancer. |
| Cardiovascular disease   | Diseases of heart; Hypertension without heart disease; Cerebrovascular diseases; Atherosclerosis; Aortic aneurysm and dissection; Other diseases of arteries, arterioles, capillaries. | 
| Communicable disease     | Tuberculosis; Syphilis; HIV (1987+), Septicemia; Other infectious and parasitic diseases; Pneumonia and influenza |
| Non-communicable disease | Diabetes Mellitus; Alzheimers (ICD9 and 10 only); Chronic obstructive pulmonary disease and allied conditions; Stomach and duodenal ulcers; Chronic liver disease and cirrhosis; Nephritis, Nephrotic Syndrome and Nephrosis.  |
| Injuries                 | Accidents and Adverse Effects; Suicide and Self-inflicted Injury; Homicide and Legal Intervention. |
| All other causes         | Complications of pregnancy, childbirth, puerperium; Congenital anomalies; Certain conditions originating in perinatal period; Symptoms, signs, and ill-defined conditions; Other cause of death. |
